Insurance choice, decision assistance, and insurer steering

### 2021-05-03 finals/creed/creed-202105.Rmd
creed slides

.center[
  ![:scale 200%](../pics/kff-uninsured.png)
]

.footnote[
  [Kaiser Family Foundation, November, 2020](https://www.kff.org/policy-watch/millions-of-uninsured-americans-are-eligible-for-free-aca-health-insurance/)
]

@@

---
# Consumer choice model

Household i's Utility for Plan j<br>

$\begin{eqnarray}U_{ij} &=& \alpha_i \underbrace{p_{ij}}_{cons.\\prem.} \hspace{0.075in} + \hspace{0.075in} \underbrace{x'_{j}}_{obs.\\prod.} \beta \hspace{0.075in} + \hspace{0.075in} \underbrace{d'_{i}}_{dem.\\char.} \varphi \hspace{0.075in} + \hspace{0.075in} \underbrace{\xi_{j}}_{unobs.\\prod.} \hspace{0.075in} + \hspace{0.075in} \underbrace{\epsilon_{ij}}_{error} \nonumber\end{eqnarray}$

where<br>

$\begin{eqnarray*}\alpha_i &=& \alpha + d'_i \gamma \end{eqnarray*}$

$\begin{eqnarray*}\underbrace{p_{ij}}_{cons.\\prem.} &=& \underbrace{r_{i}}_{rating\\ fact.} \underbrace{p_j}_{full\\prem.} \hspace{0.075in} - \hspace{0.075in} \underbrace{\max\{p_b - \mathrm{cap}_i,0\}}_{subsidy} \end{eqnarray*}$

# Consumer choice model

Utility for Outside Option (Uninsured)<br>

$\begin{eqnarray*} U_{i0} &=& \alpha_i \underbrace{\rho_i}_{\textrm{penalty}} \hspace{0.075in} + \hspace{0.075in}  \underbrace{\epsilon_{i0}}_{error} \end{eqnarray*}$

@@

- Denote by $N_{j}^{1} = \sum_{i} 1(Y_{ij}^{1}=1 | \text{Assist})$ the the count of enrollees in plan $j$ had they received decision assistance, conditional on receiving assistance

- Denote by $N_{j}^{0} = \sum_{i} 1(Y_{ij}^{0}=1 | \text{Assist})$ the count of enrollees in plan $j$ had they not received decision assistance, conditional on receiving assistance

@@

$$ATT = N_{j}^{1} - N_{j}^{0} $$

@@

1. Estimate nested logit model among those without any decision assistance, $\hat{\theta}$

2. Use $\hat{\theta}$ to form predicted counterfactual, $\hat{N}_{j}^{0} = \sum_{i} \widehat{Pr[Y_{ij}^{0} | \text{Assist}]}$

3. Use predicted value in ATT estimate: $$\underbrace{N_{j}^{1}}_{Observed} - \underbrace{\hat{N}_{j}^{0}}_{Predicted}$$

4. Aggregate to level of insurer and metal tier

>>>>

.center[
  ![:scale 75%](../pics/kff-uninsured.png)
]

@@

@@

- Denote by $Pr[Y_{ij}^{1}]$ the probability of household $i$ selecting plan $j$ with decision assistance

- Denote by $Pr[Y_{ij}^{0}]$ the probability of household $i$ selecting plan $j$ without decision assistance

@@

$$ATT = Pr[Y_{ij}^{1} | \text{Assist}] - Pr[Y_{ij}^{0} | \text{Assist}]$$

@@

1. Estimate multinomial logit model among those without any decision assistance, $\hat{\theta}$

2. Use $\hat{\theta}$ to form predicted counterfactual, $\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}$

3. Use predicted value in ATT estimate: $$\underbrace{Pr[Y_{ij}^{1} | \text{Assist}]}_{Observed} - \underbrace{\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}}_{Predicted}$$

4. Aggregate to level of insurer and metal
